---
title: "Testing correlations between sentiment analysis aggregates"
author: "Hannah Metzler & David Garcia"
output:
  pdf_document
url_colour: blue
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warnings=FALSE)
```

```{r, libraries, results='hide'}
library(tidyverse)
library(plotly)
library(lubridate)
library(cowplot)
library(gridExtra)
library(boot)
```

# Comparing LIWC and RoBERTa

For RoBERTa scores, only tweets with a likelihood of above 0.9 of expressing that emotion (as labelled by humans) are counted as hits for that emotion. 

## USA

```{r}
RobertaDF <- read.csv("data_roberta/USRoberta.csv") 
LIWCDF <- read.csv("data_main/CountriesData.csv")
LIWCDF %>% filter(country=="usa")-> LIWCDF

plotDF = RobertaDF %>% 
  rename("anxiety" = "fear") %>% #same name as the LIWC wordlist
  #transform to long format
  gather(key = "wordlist", value = "RoBERTa", "anger", "anxiety", "positive", "sadness") %>% 
  #add LIWC df
  inner_join(LIWCDF) %>% 
  rename("LIWC" = pr, 
         "emotion" = wordlist) %>% #in roberta its not a wordlist, but a predicted label
  mutate(RoBERTa = RoBERTa*100) %>% 
  mutate(date = as.Date(date))
```


```{r, scatterplot LIWC Roberta US, fig.asp = 1}
svanx = ggplot(data = filter(plotDF, emotion=="anxiety"), aes(x = LIWC, y =RoBERTa)) +
  geom_smooth( method = "lm", color="gray", alpha=0.3, se = F, linetype=2) +
  geom_point(size=2, colour="orange")+
   theme_bw()+ theme(text=element_text(size=12), axis.text=element_text(size=12), legend.title=element_blank(), legend.position="right")+ggtitle("Anxiety")+
  labs(x="LIWC")
svsad = ggplot(data = filter(plotDF, emotion=="anger"), aes(x = LIWC, y =RoBERTa)) +
  geom_smooth( method = "lm", color="gray", alpha=0.3, se = F, linetype=2) +
  geom_point(size=2, colour="red")+
   theme_bw()+ theme(text=element_text(size=12), axis.text=element_text(size=12), legend.title=element_blank(), legend.position="right")+ggtitle("Anger")
svang = ggplot(data = filter(plotDF, emotion=="sadness"), aes(x = LIWC, y =RoBERTa)) +
  geom_smooth( method = "lm", color="gray", alpha=0.3, se = F, linetype=2) +
  geom_point(size=2, colour="blue")+
   theme_bw()+ theme(text=element_text(size=12), axis.text=element_text(size=12), legend.title=element_blank(), legend.position="right")+ggtitle("Sadness")
svpos = ggplot(data = filter(plotDF, emotion=="positive"), aes(x = LIWC, y =RoBERTa)) +
  geom_smooth( method = "lm", color="gray", alpha=0.3, se = F, linetype=2) +
  geom_point(size=2, colour="cyan")+
   theme_bw()+ theme(text=element_text(size=12), axis.text=element_text(size=12), legend.title=element_blank(), legend.position="right")+ggtitle("Positive")

sv_us = cowplot::plot_grid(svanx,  svang, svsad, svpos, ncol=2, vjust = 1, align = "hv"); #labels=c("a", "b", "c", "d"), label_size=25, 
sv_us 
ggsave("./figures/sentimentvalidation_USA.pdf", plot=sv_us)
```

## Correlation tests

Both Pearson and Spearman, given that there are quite a few outliers in the plots above that likely influence the correlation
* Variable "pr" is the percentage of tweets with LIWC words (uncorrected for the baseline)
* The variable name that is an emotion (fear, anger, sadness, pos) is the RoBERTa score. 

```{r}
LIWCDF %>% filter(wordlist=="anxiety") -> df1
df1 <- inner_join(RobertaDF, df1, by="date")
cor.test(df1$pr, df1$fear, method="pearson")
cor.test(df1$pr, df1$fear, method="spearman")

LIWCDF %>% filter(wordlist=="sadness") -> df1
df1 <- inner_join(RobertaDF, df1, by="date")
cor.test(df1$pr, df1$sadness, method="pearson")
cor.test(df1$pr, df1$sadness, method="spearman")

LIWCDF %>% filter(wordlist=="anger") -> df1
df1 <- inner_join(RobertaDF, df1, by="date")
cor.test(df1$pr, df1$anger, method="pearson")
cor.test(df1$pr, df1$anger, method="spearman")

LIWCDF %>% filter(wordlist=="positive") -> df1
df1 <- inner_join(RobertaDF, df1, by="date")
cor.test(df1$pr, df1$pos, method="pearson")
cor.test(df1$pr, df1$pos, method="spearman")
```

Correlations for a paper: 

US anxiety r=0.64 t(475) = 18.1, p < 0.001
US sadness r=0.71, t(475) = 21.84, p < 0.001
US anger r=0.73, t(475) = 23.48, p < 0.001
US positive r=0.42, t(475) = 10.21, p < 0.001

### Time series

```{r}
#plot
ggplot() +
  geom_line(data = plotDF, aes(x=date, y=RoBERTa, colour=emotion))+
  geom_line(data = plotDF, aes(x=date, y=LIWC, colour=emotion), linetype=2, show.legend=T)+
  # ylab("% Difference over 2019 baseline") +  ggtitle("US state comparison")+ #plot title
  scale_colour_manual(values = c("anxiety"="orange", "anger"="red", "sadness"="blue", "positive"="cyan"))+
   theme_bw()+ theme(
                    axis.title.x = element_blank(),
                    legend.title=element_blank(), legend.position="bottom",
                    plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"))+
  scale_x_date(date_breaks="1 month", date_labels = "%b")+
  scale_y_continuous(name = "% tweets RoBERTa", sec.axis = dup_axis(name="% tweets LIWC"))
```


## Spain

```{r}
RobertaDF <- read.csv("data_roberta/ESRoberta.csv")
LIWCDF <- read.csv("data_main/CountriesData.csv")
LIWCDF %>% filter(country=="esp") -> LIWCDF

plotDF = RobertaDF %>% 
  rename("anxiety" = "fear") %>% #same name as the LIWC wordlist
  #transform to long format
  gather(key = "wordlist", value = "RoBERTa", "anger", "anxiety", "positive", "sadness") %>% 
  #add LIWC df
  inner_join(LIWCDF) %>% 
  rename("LIWC" = pr, 
         "emotion" = wordlist) %>% #in roberta its not a wordlist, but a predicted label
  mutate(RoBERTa = RoBERTa*100) %>% 
    mutate(date = as.Date(date))
```

```{r, scatterplot LIWC Roberta Spain, fig.asp = 1}

svanx = ggplot(data = filter(plotDF, emotion=="anxiety"), aes(x = LIWC, y =RoBERTa)) +
  geom_smooth( method = "lm", color="gray", alpha=0.3, se = F, linetype=2) +
  geom_point(size=2, colour="orange")+
   theme_bw()+ theme(text=element_text(size=12), axis.text=element_text(size=12), legend.title=element_blank(), legend.position="right")+ggtitle("Anxiety")
svsad = ggplot(data = filter(plotDF, emotion=="anger"), aes(x = LIWC, y =RoBERTa)) +
  geom_smooth( method = "lm", color="gray", alpha=0.3, se = F, linetype=2) +
  geom_point(size=2, colour="red")+
   theme_bw()+ theme(text=element_text(size=12), axis.text=element_text(size=12), legend.title=element_blank(), legend.position="right")+ggtitle("Anger")
svang = ggplot(data = filter(plotDF, emotion=="sadness"), aes(x = LIWC, y =RoBERTa)) +
  geom_smooth( method = "lm", color="gray", alpha=0.3, se = F, linetype=2) +
  geom_point(size=2, colour="blue")+
   theme_bw()+ theme(text=element_text(size=12), axis.text=element_text(size=12), legend.title=element_blank(), legend.position="right")+ggtitle("Sadness")
svpos = ggplot(data = filter(plotDF, emotion=="positive"), aes(x = LIWC, y =RoBERTa)) +
  geom_smooth( method = "lm", color="gray", alpha=0.3, se = F, linetype=2) +
  geom_point(size=2, colour="cyan")+
   theme_bw()+ theme(text=element_text(size=12), axis.text=element_text(size=12), legend.title=element_blank(), legend.position="right")+ggtitle("Positive")

sv_esp = cowplot::plot_grid(svanx,  svang, svsad, svpos, ncol=2,  vjust = 1, align="hv"); #labels=c("a", "b", "c", "d"), label_size=25,
sv_esp
ggsave("./figures/sentimentvalidation_spain.pdf", plot=sv_esp)
```

Correlation tests: Both Pearson and Spearman, given that there are quite a few outliers in the plots above that likely influence the correlation 

```{r}
LIWCDF %>% filter(wordlist=="anxiety") -> df1
df1 <- inner_join(RobertaDF, df1, by="date")
cor.test(df1$pr, df1$fear, method="pearson")
cor.test(df1$pr, df1$fear, method="spearman")

LIWCDF %>% filter(wordlist=="sadness") -> df1
df1 <- inner_join(RobertaDF, df1, by="date")
cor.test(df1$pr, df1$sadness, method="pearson")
cor.test(df1$pr, df1$sadness, method="spearman")

LIWCDF %>% filter(wordlist=="anger") -> df1
df1 <- inner_join(RobertaDF, df1, by="date")
cor.test(df1$pr, df1$anger, method="pearson")
cor.test(df1$pr, df1$anger, method="spearman")

LIWCDF %>% filter(wordlist=="positive") -> df1
df1 <- inner_join(RobertaDF, df1, by="date")
cor.test(df1$pr, df1$pos, method="pearson")
cor.test(df1$pr, df1$pos, method="spearman")
```

### Time series

```{r}
#plot
ggplot() +
  geom_line(data = plotDF, aes(x=date, y=RoBERTa, colour=emotion))+
  geom_line(data = plotDF, aes(x=date, y=LIWC, colour=emotion), linetype=2, show.legend=T)+
  # ylab("% Difference over 2019 baseline") +  ggtitle("US state comparison")+ #plot title
  scale_colour_manual(values = c("anxiety"="orange", "anger"="red", "sadness"="blue", "positive"="cyan"))+
   theme_bw()+ theme(
                    axis.title.x = element_blank(),
                    legend.title=element_blank(), legend.position="bottom",
                    plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"))+
  scale_x_date(date_breaks="1 month", date_labels = "%b")+
  scale_y_continuous(name = "% tweets RoBERTa", sec.axis = dup_axis(name="% tweets LIWC"))
```

Correlations for a paper: 

Spain anxiety r=0.37, t(472) = 8.75, p < 0.001
Spain sadness r=0.42, t(472) = 10.06, p < 0.001
Spain anger r=0.70, t(472) = 21.07, p < 0.001
Spain positive r=0.64, t(472) = 18.22, p < 0.001

# Roberta for different follower thresholds

## US

### Correlations of RoBERTa in sample of tweets excluding above 50.000 or 100.000 followers, with 5000 followers

```{r}
RobertaDF <- read.csv("data_roberta/USRoberta.csv")
RobertaDF5K <- read.csv("data_roberta/USRoberta-5K.csv")
cor.test(RobertaDF$anger, RobertaDF5K$anger)
cor.test(RobertaDF$fear, RobertaDF5K$fear)
cor.test(RobertaDF$positive, RobertaDF5K$positive)
cor.test(RobertaDF$sadness, RobertaDF5K$sadness)
```

### Correlations of RoBERTa in sample of tweets excluding above 50.000 or 100.000 followers, with 2000 followers

```{r}
RobertaDF2K <- read.csv("data_roberta/USRoberta-2K.csv")
cor.test(RobertaDF$anger, RobertaDF2K$anger)
cor.test(RobertaDF$fear, RobertaDF2K$fear)
cor.test(RobertaDF$positive, RobertaDF2K$positive)
cor.test(RobertaDF$sadness, RobertaDF2K$sadness)
```

## Spain

### Correlations of RoBERTa in sample of tweets excluding above 50.000 or 100.000 followers, with 5000 followers

```{r}
RobertaDF <- read.csv("data_roberta/ESRoberta.csv")
RobertaDF5K <- read.csv("data_roberta/ESRoberta-5K.csv")
cor.test(RobertaDF$anger, RobertaDF5K$anger)
cor.test(RobertaDF$fear, RobertaDF5K$fear)
cor.test(RobertaDF$positive, RobertaDF5K$positive)
cor.test(RobertaDF$sadness, RobertaDF5K$sadness)
```

### Correlations of RoBERTa in sample of tweets excluding above 50.000 or 100.000 followers, with 2000 followers

```{r}
RobertaDF2K <- read.csv("data_roberta/ESRoberta-2K.csv")
cor.test(RobertaDF$anger, RobertaDF2K$anger)
cor.test(RobertaDF$fear, RobertaDF2K$fear)
cor.test(RobertaDF$positive, RobertaDF2K$positive)
cor.test(RobertaDF$sadness, RobertaDF2K$sadness)
```



