---
title: "Collective Emotions during the Covid-19 Outbreak"
subtitle: "<span style='font-size: 27px'>Supplementary Information 2: Word frequencies in tweets </style>"
author: "Hannah Metzler, Bernard Rimé, Max Pellert, Thomas Niederkrotenthaler, Anna Di Natale & David Garcia"
output:
     html_document: default
     
    #to get a pdf of this markdown: print to html, and then print as pdf from the browser (otherwise the emoji don't work)
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warnings=FALSE, comment = "")
```

```{r, libraries, results='hide'}
library(tidyverse)
library(rjson)
library(patchwork)
library(cowplot)
library(kableExtra)
```

## Methods

We use word frequency plots of the 60 most frequent terms to check that there are no strange outlier words that drive the changes we observe. If tweets repeatedly using an LIWC term in the wrong context were very frequent, other non-emotional words from these tweets would indicate this problem. (Example: high frequencies of the term "lov*" together with "McDonalds" would indicate that the sentence "i am lovin it, McDonalds" influenced our tweet count for positive emotions.) Therefore, we used word frequency plots on all words, rather than just LIWC terms. 

We made one plot per language, emotion and time period, including the baseline, and the outbreak period of 5 weeks after the day with 30 cases. The baseline period plots provide an overall idea of what each dictionary picks up on in general, independent from a particular event, while plots for the outbreak period show which particular words became more frequent in relation to Covid-19.<!-- (- We could use the control period to check if the year 2020 is comparable to the baseline, to see the more immediate change in words from pre to post covid in 2020.) --> We chose the following example countries for each language: 

* German: Germany
* English: United Kingdom
* Dutch: Netherlands
* French: France
* Italian: Italy
* Spanish: Spain

We inspected which *topics* people were emotional about (i.e. non-LIWC terms that occur frequently), and which of the *LIWC terms* were most frequent, i.e. importantly adding to the number of emotional tweets. Here, disproportionally large contributions by single words would indicate problems, such as one single event strongly impacting this number, without necessarily reflecting a general change in emotions. __For the countries in which more than one LIWC term made it into the top 60 terms (UK, Germany, Spain and France), we therefore report the percentage of tweets containing each term in Tables S2.1 to S2.4. We further report the frequency of the least frequent terms in the top 60 per period and emotion, as upper limits for all other LIWC terms for all six countries.__

An example for one word driving an emotion peak, from outside the analyzed time periods, is the word "terror". It accounts for a large propotion of anxiety and anger spikes after the attack in Hanau in German and Austrian tweets, and this seems to mostly reflect reports about the attack rather than a broader change in emotional expressions on Twitter. In contrast, if most words only occur in a few percent of tweets, this indicates that LIWC words that might have misfired during specific events or in a context that does not reflect the emotino did not strongly bias our emotion measures. 

<!-- In the script below, all text and code parts calculating and plotting the frequency difference between the baseline and the post-outbreak period are commented. Mark them and press Ctrl+Shift+C to uncomment them and execute the code and write the text.  -->
<!-- Finally, we make one plot for the difference in the percent of tweets that contained a particular word. Because we only have the 60 most frequent words in each category, these plots only include words that occur both during the baseline and during the outbreak. An additional plot shows the words that were part of the 60 most frequent ones during the outbreak, but were not that frequent during the baseline. -->

<!-- ## Settings -->

```{r, settings, cache=T}
#general settings for all plots: Unicode for all emojis/symbols (these can only be printed to html)
emoji_joy_tears =  intToUtf8(0x1f602)
emoji_heart_eyes =  intToUtf8(0x1f600 + 13)
emoji_crying =  intToUtf8(0x1f600 + 45)
emoji_angry =  intToUtf8(0x1f600 + 32)
emoji_rolling = intToUtf8(0x0001f923)
emoji_smiling_eyes = intToUtf8(0x1f600 + 10)
thumbs_up = intToUtf8(0x1F44D)
hugging_face = intToUtf8(0x1F917)#hugging face
emoji_heart_eyes = intToUtf8(0x1f600 + 13)#smiling face with heart-shaped eyes
emoji_winking = intToUtf8(0x1f600 + 9)#winking face
emoji__kiss=intToUtf8(0x1f600 + 24) #face throwing a kiss
emoji_thinking = intToUtf8(0x1F914)#thinking face
emoji_grinning = intToUtf8(0x1F601)# grinning face with smiling eyes
emoji_sweat = intToUtf8(0x1F605)#smiling face with open mouth and cold sweat
clover = intToUtf8(0x1F340)#four leaf clover
emoji_rolling_eyes = intToUtf8(0x1F644)#face with rolling eyes
hot_beverage = intToUtf8(0x2615)
emoji_sunglass = intToUtf8(0x1F60E)
see_no_monkey = intToUtf8(0x1F648)
pointing_index=intToUtf8(0x1F449)
blue_circle=intToUtf8(0x1F535)
clapping = intToUtf8(0x1F44F)
biceps = intToUtf8(0x1F4AA)
black_heart = intToUtf8(0x1F5A4)
delicious = intToUtf8(0x1F60B)
pleading = intToUtf8(0x1F97A)

#function to recode text in terms into emoji:
makeemoji = function(term){
   recode(term, "face with tears of joy" = emoji_joy_tears, "rolling on the floor laughing" = emoji_rolling,
                       "loudly crying face" = emoji_crying, "smiling face with smiling eyes" = emoji_smiling_eyes, "thumbs up sign" = thumbs_up,
                       "hugging face" = hugging_face, "smiling face with heart-shaped eyes"=emoji_heart_eyes, "winking face"=emoji_winking,
                       "face throwing a kiss"=emoji__kiss, "thinking face" = emoji_thinking, "grinning face with smiling eyes"=emoji_grinning,
                       "smiling face with open mouth and cold sweat"=emoji_sweat,"four leaf clover"=clover,
                       "face with rolling eyes"=emoji_rolling_eyes, "hot beverage" = hot_beverage, "smiling face with sunglasses"=emoji_sunglass, "see-no-evil monkey" = see_no_monkey, "white right pointing backhand index" = pointing_index, "large blue circle" =blue_circle, "clapping hands sign"=clapping, "flexed biceps"=biceps, "heavy black heart"=black_heart, "face savouring delicious food"=delicious, "pleading face"=pleading)
}

#show words that overlap between baseline and outbreak period: True or false
#if only terms that are not in total tweets should be shown (for baseline and 5 week plots), set to false (no overlapping terms)
overlap = TRUE # change this in the entire script using replacement, so the cache=T statement catches that the code has changed, otherwise it just uses saved figures from the cache folder

```



\newpage

## Results 

### English: United Kingdom

```{r, prep data UK, cache=T}
overlap = TRUE
country <- "uk"
totalbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_total_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

angerbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anger_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq))

sadbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_sad_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>%filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq))

anxbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anxiety_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>%filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

posbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_positive_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

# 5 week after outbreak
totaldf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_total_20200301-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  # exchange emojji picture for description
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# difference to baseline
totaldiffdf = totalbasedf %>%  select(-Type) %>%
  right_join(totaldf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))

angerdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anger_20200301-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
    mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# add columns for terms that exist in total and anger (36) and calculate difference
# angerdiffdf = angerdf %>%  select(-Type) %>%
#   inner_join(angerdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
angerdiffdf = angerbasedf %>%  select(-Type) %>%
  right_join(angerdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
#keep only terms that are not in totaldf
if(overlap==FALSE){angerdf = angerdf %>%
  filter(!term %in% totaldf$term)}

saddf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_sad_20200301-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# saddiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(saddf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
saddiffdf = sadbasedf %>%  select(-Type) %>%
  right_join(saddf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){saddf = saddf %>%
  filter(!term %in% totaldf$term)}

anxdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anxiety_20200301-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
   mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# anxdiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(anxdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
anxdiffdf = anxbasedf %>%  select(-Type) %>%
  right_join(anxdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){anxdf = anxdf %>%
  filter(!term %in% totaldf$term)}

posdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_positive_20200301-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
   mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# posdiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(posdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
posdiffdf = posbasedf %>%  select(-Type) %>%
  right_join(posdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){posdf = posdf %>%
  filter(!term %in% totaldf$term)}


### how many terms occur in more than 1% of the tweets (during outbreak)?
country = rep(country, nrow(anxdf)+nrow(angerdf)+nrow(saddf)+nrow(posdf))
emotion = rep(c("anxiety", "anger", "sadness", "positive"), times = c(nrow(anxdf),nrow(angerdf),nrow(saddf),nrow(posdf)))
alldf = cbind(rbind(anxdf, angerdf, saddf, posdf), emotion, country)

```

<!-- If overlap is set to FALSE: For tweets that contain emotional terms, the plot only shows terms not included among the most frequent in total tweet volume. -->

```{r, plots baseline UK,  fig.width=12, fig.height=9, cache=T, fig.cap ="**Figure S2.1. Most frequent terms in tweets per category during the baseline in the UK.** Percentage of tweets with the term."}
overlap = TRUE
ymax=8
ptot <- ggplot(totalbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="gray") +
  geom_text(aes(label = term, y=freq+0.1*nchar(as.character(term))/2), vjust = 0.5) +  ylim(0,ymax)+ylab("% total tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank()) 

panx <- ggplot(anxbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="orange") +
  geom_text(aes(label = term, y=freq+0.1*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% anxiety tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

psad <- ggplot(sadbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="blue") +
  geom_text(aes(label = term, y=freq+0.08*nchar(as.character(term))), vjust = 0.5)+ ylim(0,ymax)+ ylab("% sadness tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

pang <- ggplot(angerbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="red") +
  geom_text(aes(label = term, y=freq+0.1*nchar(as.character(term))/2), vjust = 0.5)+ylim(0,ymax)+ ylab("% anger tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ppos <- ggplot(posbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="cyan") +
  geom_text(aes(label = term, y=freq+0.1*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% positive tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ptot | panx | psad | pang |  ppos
```

\newpage

<!-- If overlap is set to FALSE: For tweets that contain emotional terms, the plot only shows terms not included among the most frequent in total tweet volume. -->

```{r, plots outbreak UK,  fig.width=12, fig.height=9, cache=T, fig.cap="**Figure S2.2. Most frequent terms in tweets per category during the 5 weeks after the outbreak in the UK.** Percentage of tweets with the term."}
overlap = TRUE
ymax = 13.4
ptot <- ggplot(totaldf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="gray") +
  geom_text(aes(label = term, y=freq+0.3*nchar(as.character(term))/2), vjust = 0.5) +  ylim(0,ymax)+ylab("% total tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank()) 

panx <- ggplot(anxdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="orange") +
  geom_text(aes(label = term, y=freq+0.15*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% anxiety tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

psad <- ggplot(saddf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="blue") +
  geom_text(aes(label = term, y=freq+0.15*nchar(as.character(term))), vjust = 0.5)+ ylim(0,ymax)+ ylab("% sadness tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

pang <- ggplot(angerdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="red") +
  geom_text(aes(label = term, y=freq+0.3*nchar(as.character(term))/2), vjust = 0.5)+ylim(0,ymax)+ ylab("% anger tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ppos <- ggplot(posdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="cyan") +
  geom_text(aes(label = term, y=freq+0.3*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% positive tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())
ptot | panx | psad | pang | ppos
```


<!-- ### Difference: Volume of tweets with the term during baseline vs. 5 outbreak weeks -->

<!-- What increases, what decreases after the outbreak? Only words that are among the 60 most frequent during baseline are shown. New terms are shown in the second row. -->

<!-- ```{r, plots difference UK,  fig.width=12, fig.height=7} -->
<!-- overlap = TRUE -->
<!-- ymax = 5; ymin = -0.7 -->

<!-- ptot <- ggplot(filter(totaldiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="grey") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.1*nchar(as.character(term))/2),vjust = 0.5)+ ylim(ymin,ymax)+ ylab("total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ ggtitle("Outbreak vs. baseline") -->

<!-- panx <- ggplot(filter(anxdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="orange") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.05*nchar(as.character(term))/2),vjust = 0.5)+ ylim(ymin,ymax)+ ylab("anxiety tweets vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank()) -->
<!-- psad <- ggplot(filter(saddiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="blue") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.03*nchar(as.character(term))), vjust = 0.5)+  ylim(ymin,ymax)+ ylab("sadness vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- pang <- ggplot(filter(angerdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="red") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.05*nchar(as.character(term))/2), vjust = 0.5)+ ylim(ymin,ymax)+ ylab("anger vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- ppos <- ggplot(filter(posdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="cyan") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.05*nchar(as.character(term))/2), vjust = 0.5)+  ylim(ymin,ymax)+ ylab("positive vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- ptot | panx | psad | pang |ppos -->
<!-- ``` -->


<!-- ```{r  fig.width=12, fig.height=5} -->
<!-- #plots with new terms -->
<!-- ymax = 6 -->
<!-- newtot <- ggplot(filter(totaldiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="grey")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.03*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+ ylab("anxiety tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("New terms during outbreak") -->
<!-- newanx <- ggplot(filter(anxdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="orange")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.03*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+ ylab("anxiety tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank()) -->
<!-- newsad <- ggplot(filter(saddiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="blue")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.01*nchar(as.character(term))/2),vjust = 0.5)+  ylim(0,ymax)+ ylab("sadness tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newang <- ggplot(filter(angerdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="red")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.01*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+  ylab("anger tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newpos <- ggplot(filter(posdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="cyan")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.01*nchar(as.character(term))/2),vjust = 0.5)+  ylim(0,ymax)+ ylab("positive tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newtot | newanx | newsad | newang |  newpos -->
<!-- ``` -->

```{r, english liwc terms, cache = TRUE}
#adapt wordlist dfs for a table
make_df_for_tab =function(df, period){
  #df as emotion name
  emo = as.character(substitute(df))
  df1 = df %>% 
    select(-c(Type, Mentions)) %>% 
    rename(Term = term, Percent = freq) %>% 
    mutate(Period = rep(period, nrow(df)), 
           Emotion = rep(str_replace(emo, "basedf|df", ""), nrow(df)))
  return(df1)
}
angb = make_df_for_tab(angerbasedf, "baseline")
anxb = make_df_for_tab(anxbasedf, "baseline")
sadb = make_df_for_tab(sadbasedf, "baseline")
posb = make_df_for_tab(posbasedf, "baseline")
ang = make_df_for_tab(angerdf, "outbreak")
anx = make_df_for_tab(anxdf, "outbreak")
sad = make_df_for_tab(saddf, "outbreak")
pos = make_df_for_tab(posdf, "outbreak")

#get and preprocess LIWC dictionaries
#directory with the dictionary files
dirdictionaries <- '~/Documents/dictionaries/'
#load dictionaries and delete the *, because it means zero or more matches of the previous letter. 
#$ at end of each liwc term without * searches only for words exactly ending like the liwc term 
dictang <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'english2007_anger.txt'))$V1, "$"), "\\*\\$", "")
dictanx <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'english2007_anx.txt'))$V1, "$"), "\\*\\$", "")
dictsad <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'english2007_sad.txt'))$V1, "$"), "\\*\\$", "")
dictpos <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'english2007_posemo_noambiguous.txt'))$V1, "$"), "\\*\\$", "")

# ## add "[\\s-]|^" at the start of each word to also include matches that include "-" or a white space (e.g. self-isolating, tut mir leid), and so that words have to start with the search term otherwise (^) (so "ok" does not match with "book")
dictang = paste0("[\\s-]", dictang, "|^", dictang)
dictanx = paste0("[\\s-]", dictanx, "|^", dictanx)
dictsad = paste0("[\\s-]", dictsad, "|^", dictsad)
dictpos = paste0("[\\s-]", dictpos, "|^", dictpos)

#add escape character to smiley brackets and colon
dictpos <- str_replace_all(dictpos, "\\(", "\\\\\\(")
dictpos <- str_replace_all(dictpos, "\\)", "\\\\\\)")
dictpos <- str_replace_all(dictpos, "\\:", "\\\\\\:")

# filter out LIWC terms
angbliwc =  angb %>% 
  filter(str_detect(Term, paste(dictang, collapse = "|")))
anxbliwc =  anxb %>% 
  filter(str_detect(Term, paste(dictanx, collapse = "|")))
sadbliwc =  sadb %>% 
  filter(str_detect(Term, paste(dictsad, collapse = "|")))
posbliwc =  posb %>% 
  filter(str_detect(Term, paste(dictpos, collapse = "|")))
angliwc =  ang %>% 
  filter(str_detect(Term, paste(dictang, collapse = "|")))
anxliwc =  anx %>% 
  filter(str_detect(Term, paste(dictanx, collapse = "|")))
sadliwc =  sad %>% 
  filter(str_detect(Term, paste(dictsad, collapse = "|")))
posliwc =  pos %>% 
  filter(str_detect(Term, paste(dictpos, collapse = "|")))

liwc = rbind(angbliwc, anxbliwc, sadbliwc, posbliwc, angliwc, anxliwc, sadliwc, posliwc) %>% 
  mutate(Emotion = recode(Emotion, pos = "positive", anx = "anxiety", sad = "sadness", ang = "anger")) %>% 
  relocate(c("Period", "Emotion"), .before="Term")
liwc_uk = liwc
```

**Table S2.1. Frequency of LIWC terms among the 60 most frequent terms in the UK per time period and emotion.**

```{r}
knitr::kable(liwc,
          digits = 2,
          format="html",
           table.attr = "style='width:50%;'") %>% 
  kable_classic(full_width=T, position = "left",  font_size = 13) 
```

```{r}
mins =round((c(min(anxb$Percent), min(anx$Percent), min(angb$Percent), min(ang$Percent), min(sadb$Percent), min(sad$Percent), min(posb$Percent), min(pos$Percent))),2)
lowest_percent = paste(mins[1], "and", mins[2], "for anxiety,", mins[3], "and", mins[4], "for anger,",mins[5], "and", mins[6], "for sadness, and",mins[7], "and", mins[8], "for positive emotions")
```

**Observations**

*LIWC words*: 

__No LIWC term was present in more than `r round(max(liwc$Percent), digits = 2)`% of tweets expressing a particular emotion. During the outbreak, panic buying and self-isolating were the two most frequent anxiety and sadness terms from LIWC. Percent of tweets containing all LIWC terms not listed in Table S2.1. during baseline and outbreak, respectively, were below `r lowest_percent` (percentages for words at rank 60 per dictionary and time period).__

*Topics*: 

* General: More tweets mentioned people, home, work and time during the outbreak vs. baseline. The word social became part of the 60 most frequent terms, and occurs often in anxiety and sadness tweets, possibly linked to social distancing. 
* The increase in mentions of "people" was stronger for negative emotional tweets than tweets in general. Mentions of home increased most in anxiety and sadness tweets.
* Shift in topic from political to pandemic: In 2019, tweets, and in particular emotional ones, often mentione political issues related to Brexit, the EU and the Labour party (leave, left, party, UK, vote,...). During the outbreak, these words are gone, and instead words like coronavirus, health, staff, #covid-19, virus are much more frequent. 
* "Government" is more frequent in anxious and angry emotional than other tweets. 
* Anxiety: The word health is one of the most frequent. Coronavirus ranks higher than in other emotion categories. People worry about panic buying, and about (hospital) staff. Positive words like good, life, great and smiley were less frequent in anxiety tweets than during baseline. 
* Sadness: Sadness tweets evolve around staying at home, the famliy, the government, support, social, suggesting staying at home as a factor for sadness. 
* Anger tweets mentioned the government more than during baseline. 
* Positive emotions: The word lockdown ranks higher than in other tweets. People seem to have noticed positive aspects of staying at home. Other new highly frequent words during the outbreak: online, books, family, team. 

\newpage


### German: Germany

```{r, prep data Germany}
overlap = TRUE
country <- "germany"
totalbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_total_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term!="unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

angerbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anger_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>%
 mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

sadbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_sad_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>%
mutate(term = makeemoji(term))%>%
  arrange(desc(freq)) %>%
  #remove misparsed Co, part of abbreviated twitter links like https://t.co/i9Ck6mMTw1
  filter(term!="Co")

anxbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anxiety_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>%
mutate(term = makeemoji(term))%>%arrange(desc(freq))

posbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_positive_2019.csv"), header=T, skip=6) %>% 
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term!="unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

country="germany"
totaldf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_total_20200227-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  # exchange emojji picture for description
    mutate(term = makeemoji(term))%>%  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# difference to baseline
totaldiffdf = totalbasedf %>%  select(-Type) %>%
  right_join(totaldf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, 
         term = tolower(term))

angerdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anger_20200227-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
    mutate(term = makeemoji(term))%>%  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# add columns for terms that exist in total and anger (36) and calculate difference
# angerdiffdf = angerdf %>%  select(-Type) %>%
#   inner_join(angerdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
angerdiffdf = angerbasedf %>%  select(-Type) %>%
  right_join(angerdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base,  term = tolower(term))
if(overlap==FALSE){angerdf = angerdf %>%
  filter(!term %in% totaldf$term)}

saddf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_sad_20200227-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
   mutate(term = makeemoji(term))%>%  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# saddiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(saddf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
saddiffdf = sadbasedf %>%  select(-Type) %>%
  right_join(saddf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){saddf = saddf %>%
  filter(!term %in% totaldf$term)}

anxdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anxiety_20200227-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# anxdiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(anxdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
anxdiffdf = anxbasedf %>%  select(-Type) %>%
  right_join(anxdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){anxdf = anxdf %>%
  filter(!term %in% totaldf$term)}

posdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_positive_20200227-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
   mutate(term = makeemoji(term))%>%  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# posdiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(posdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
posdiffdf = posbasedf %>%  select(-Type) %>%
  right_join(posdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){posdf = posdf %>%
  filter(!term %in% totaldf$term)}

### how many terms occur in more than 1% of the tweets (during outbreak)? 
country = rep(country, nrow(anxdf)+nrow(angerdf)+nrow(saddf)+nrow(posdf))
emotion = rep(c("anxiety", "anger", "sadness", "positive"), times = c(nrow(anxdf),nrow(angerdf),nrow(saddf),nrow(posdf))) 
countrydf = cbind(rbind(anxdf, angerdf, saddf, posdf), emotion, country)
alldf = rbind(alldf, countrydf)
```


<!-- For tweets that contain emotional terms, the plot only shows terms not included among the most frequent in total tweet volume. -->

```{r  fig.width=12, fig.height=9, cache=T, fig.cap ="**Figure S2.3. Most frequent terms in tweets per category during the baseline in Germany.** Percentage of tweets with the term."}
overlap=TRUE
ymax=4
ptot <- ggplot(totalbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="gray") +
  geom_text(aes(label = term, y=freq+0.01*nchar(as.character(term))/2), vjust = 0.5) +  ylim(0,ymax)+ylab("% total tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank()) 

panx <- ggplot(anxbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="orange") +
  geom_text(aes(label = term, y=freq+0.0014*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% anxiety tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

psad <- ggplot(sadbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="blue") +
  geom_text(aes(label = term, y=freq+0.005*nchar(as.character(term))), vjust = 0.5)+ ylim(0,ymax)+ ylab("% sadness tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

pang <- ggplot(angerbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="red") +
  geom_text(aes(label = term, y=freq+0.0014*nchar(as.character(term))/2), vjust = 0.5)+ylim(0,ymax)+ ylab("% anger tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ppos <- ggplot(posbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="cyan") +
  geom_text(aes(label = term, y=freq+0.01*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% positive tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ptot | panx | psad | pang |  ppos
```

\newpage

<!-- For tweets that contain emotional terms, the plot only shows terms not included among the most frequent in total tweet volume. -->

```{r  fig.width=12, fig.height=9, cache=T, fig.cap ="**Figure S2.4. Most frequent terms in tweets per category during the 5 weeks after the outbreak in Germany.** Percentage of tweets with the term."}
overlap = TRUE
ymax = 9.1
ptot <- ggplot(totaldf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="gray") +
  geom_text(aes(label = term, y=freq+0.1*nchar(as.character(term))/2), vjust = 0.5) +  ylim(0,ymax)+ylab("% total tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank()) 

panx <- ggplot(anxdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="orange") +
  geom_text(aes(label = term, y=freq+0.2*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% anxiety tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

psad <- ggplot(saddf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="blue") +
  geom_text(aes(label = term, y=freq+0.1*nchar(as.character(term))), vjust = 0.5)+ ylim(0,ymax)+ ylab("% sadness tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

pang <- ggplot(angerdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="red") +
  geom_text(aes(label = term, y=freq+0.2*nchar(as.character(term))/2), vjust = 0.5)+ylim(0,ymax)+ ylab("% anger tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ppos <- ggplot(posdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="cyan") +
  geom_text(aes(label = term, y=freq+0.2*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% positive tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())
ptot | panx | psad | pang | ppos
```

<!-- ### Difference: Volume of tweets with the term during baseline vs. 5 outbreak weeks -->

<!-- What increases, what decreases after the outbreak? Only words that are among the 60 most frequent during baseline are shown. New terms are shown in the second row. -->

<!-- ```{r  fig.width=12, fig.height=9, cache=T} -->
<!-- ymax = 1; ymin = -0.7 -->
<!-- ptot <- ggplot(filter(totaldiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="grey") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.05*nchar(as.character(term))/2),vjust = 0.5)+ ylim(ymin,ymax)+ ylab("total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ ggtitle("Outbreak vs. baseline") -->
<!-- panx <- ggplot(filter(anxdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="orange") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.03*nchar(as.character(term))/2),vjust = 0.5)+ ylim(ymin,ymax)+ ylab("anxiety vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank()) -->
<!-- psad <- ggplot(filter(saddiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="blue") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.01*nchar(as.character(term))), vjust = 0.5)+  ylim(ymin,ymax)+ ylab("sadness vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- pang <- ggplot(filter(angerdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="red") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.03*nchar(as.character(term))/2), vjust = 0.5)+ ylim(ymin,ymax)+ ylab("anger vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- ppos <- ggplot(filter(posdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="cyan") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.03*nchar(as.character(term))/2), vjust = 0.5)+  ylim(ymin,ymax)+ ylab("positive vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- ptot |panx | psad | pang |ppos -->
<!-- ``` -->

<!-- ```{r  fig.width=12, fig.height=5} -->
<!-- #plots with new terms -->
<!-- ymax = 9 -->
<!-- newtot <- ggplot(filter(totaldiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="grey")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+ ylab("anxiety tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("New terms during outbreak") -->
<!-- newanx <- ggplot(filter(anxdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="orange")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.07*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+ ylab("anxiety tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank()) -->
<!-- newsad <- ggplot(filter(saddiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="blue")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.1*nchar(as.character(term))/2),vjust = 0.5)+  ylim(0,ymax)+ ylab("sadness tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newang <- ggplot(filter(angerdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="red")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.15*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+  ylab("anger tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newpos <- ggplot(filter(posdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="cyan")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.2*nchar(as.character(term))/2),vjust = 0.5)+  ylim(0,ymax)+ ylab("positive tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newtot | newanx | newsad | newang |  newpos -->
<!-- ``` -->



```{r, german liwc terms, cache = TRUE}
#adapt wordlist dfs for a table
angb = make_df_for_tab(angerbasedf, "baseline")
anxb = make_df_for_tab(anxbasedf, "baseline")
sadb = make_df_for_tab(sadbasedf, "baseline")
posb = make_df_for_tab(posbasedf, "baseline")
ang = make_df_for_tab(angerdf, "outbreak")
anx = make_df_for_tab(anxdf, "outbreak")
sad = make_df_for_tab(saddf, "outbreak")
pos = make_df_for_tab(posdf, "outbreak")

#get and preprocess LIWC dictionaries
#directory with the dictionary files
dirdictionaries <- '~/Documents/dictionaries/'
#load dictionaries and delete the *, because it means zero or more matches of the previous letter. 
dictang <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'german2007_anger.txt'))$V1, "$"), "\\*\\$", "")
#allow qual* only to match qualen, not e.g. qualität (mismatch)
dictang = str_replace(dictang, "qual", "qual$")
dictang = c(dictang, "qualen")
dictanx <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'german2007_anx.txt'))$V1, "$"), "\\*\\$", "")
#allow qual* only to match qualen, not e.g. qualität (mismatch)
dictanx = str_replace(dictanx, "qual", "qual$") #exclude anything except the exact match of "qual"
dictanx = c(dictanx, "qualen") #add qualen to the dictionary again as specific term
dictsad <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'german2007_sad_noambiguous.txt'))$V1, "$"), "\\*\\$", "")
dictpos <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'german2007_posemo_noambiguous.txt'))$V1, "$"), "\\*\\$", "")

#  add "[\\s-]|^" at the start of each word to also include matches that include "-" or a white space (e.g. self-isolating, tut mir leid), and so that words have to start with the search term otherwise (^) (so "ok" does not match with "book")
dictang = paste0("[\\s-]", dictang, "|^", dictang)
dictanx = paste0("[\\s-]", dictanx, "|^", dictanx)
dictsad = paste0("[\\s-]", dictsad, "|^", dictsad)
dictpos = paste0("[\\s-]", dictpos, "|^", dictpos)

#add escape character to smiley brackets and colon
dictpos <- str_replace_all(dictpos, "\\(", "\\\\\\(")
dictpos <- str_replace_all(dictpos, "\\)", "\\\\\\)")
dictpos <- str_replace_all(dictpos, "\\:", "\\\\\\:")

# filter out LIWC terms
angbliwc =  angb %>% 
  filter(str_detect(Term, paste(dictang, collapse = "|")))
anxbliwc =  anxb %>% 
  filter(str_detect(Term, paste(dictanx, collapse = "|")))
sadbliwc =  sadb %>% 
  filter(str_detect(Term, paste(dictsad, collapse = "|")))
posbliwc =  posb %>% 
  filter(str_detect(Term, paste(dictpos, collapse = "|")))
angliwc =  ang %>% 
  filter(str_detect(Term, paste(dictang, collapse = "|")))
anxliwc =  anx %>% 
  filter(str_detect(Term, paste(dictanx, collapse = "|")))
sadliwc =  sad %>% 
  filter(str_detect(Term, paste(dictsad, collapse = "|")))
posliwc =  pos %>% 
  filter(str_detect(Term, paste(dictpos, collapse = "|")))

liwc = rbind(angbliwc, anxbliwc, sadbliwc, posbliwc, angliwc, anxliwc, sadliwc, posliwc) %>% 
  mutate(Emotion = recode(Emotion, pos = "positive", anx = "anxiety", sad = "sadness", ang = "anger")) %>% 
  relocate(c("Period", "Emotion"), .before="Term")

liwc_ger = liwc
```

**Table S2.2. Frequency of LIWC terms among the 60 most frequent terms in Germany per time period and emotion.**

```{r, out.width=0.5}
knitr::kable(liwc,
          digits = 2,
          format="html",
           table.attr = "style='width:50%;'") %>% 
  kable_classic(full_width=T, position = "left",  font_size = 13) 
```

```{r}
mins =round((c(min(anxb$Percent), min(anx$Percent), min(angb$Percent), min(ang$Percent), min(sadb$Percent), min(sad$Percent), min(posb$Percent), min(pos$Percent))),2)
lowest_percent = paste(mins[1], "and", mins[2], "for anxiety,", mins[3], "and", mins[4], "for anger,",mins[5], "and", mins[6], "for sadness, and",mins[7], "and", mins[8], "for positive emotions")
```

**Observations**

*LIWC words*: 

__No LIWC term was present in more than `r round(max(liwc$Percent), digits = 2)`% of tweets expressing a particular emotion. Percent of tweets containing all LIWC terms not listed in Table S2.2. during baseline and outbreak, respectively, were below `r lowest_percent` (percentages for words at rank 60 per dictionary and time period).__

* Most terms similarly occured during the baseline and outbreak period. 
* Anxiety: Fürchte
* Sadness: Tut mir leid, enttäuscht
* Anger: 
    - Baseline: Störung, Täter, kämpfen, lächerlich, zerstört, stört
    - Outbreak: Störung, böse, kämpfen, Täter, lächerlich
* Positive: 
  - Baseline: schönen (Tag), gerne, smiley, dankeschön, glück, glückwunsch 
  - Outbreak: smiley, gerne, schönen, Glück, dankeschön

*Topics*:

* Baseline: Politics - AfD, particularly in anxiety and anger tweets. Also SPD, CDU, Politik, Frauen, Geld, Deutschland, Berlin, EU. (less politics in positive tweets)
* Outbreak: Coronavirus, Krise, Leute, Kinder, Frauen, bleiben
* Shift: More Leute, Leben, Corona, Krise, Kinder, 
* Anxiety: Qualität, Bitte, bleibt became more (Bitte bleibt zu Hause?)
* Anger: "deutschen" became less, more kämpfen
* Sadness: bleiben, zu Hause, leben, Krise, Kinder
* Positive:  bleiben, bleibt, gesund

\newpage

### Spanish: Spain

```{r, prep data Spain, cache=T}
overlap = TRUE
country <- "spain"
totalbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_total_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

angerbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anger_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq))

sadbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_sad_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>%filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq))

anxbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anxiety_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>%filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

posbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_positive_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

# 5 week after outbreak
totaldf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_total_20200228-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  # exchange emojji picture for description
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# difference to baseline
totaldiffdf = totalbasedf %>%  select(-Type) %>%
  right_join(totaldf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))

angerdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anger_20200228-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
    mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# add columns for terms that exist in total and anger (36) and calculate difference
# angerdiffdf = angerdf %>%  select(-Type) %>%
#   inner_join(angerdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
angerdiffdf = angerbasedf %>%  select(-Type) %>%
  right_join(angerdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
#keep only terms that are not in totaldf
if(overlap==FALSE){angerdf = angerdf %>%
  filter(!term %in% totaldf$term)}

saddf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_sad_20200228-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# saddiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(saddf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
saddiffdf = sadbasedf %>%  select(-Type) %>%
  right_join(saddf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){saddf = saddf %>%
  filter(!term %in% totaldf$term)}

anxdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anxiety_20200228-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
   mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# anxdiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(anxdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
anxdiffdf = anxbasedf %>%  select(-Type) %>%
  right_join(anxdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){anxdf = anxdf %>%
  filter(!term %in% totaldf$term)}

posdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_positive_20200228-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
   mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# posdiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(posdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
posdiffdf = posbasedf %>%  select(-Type) %>%
  right_join(posdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){posdf = posdf %>%
  filter(!term %in% totaldf$term)}

### how many terms occur in more than 1% of the tweets (during outbreak)?
country = rep(country, nrow(anxdf)+nrow(angerdf)+nrow(saddf)+nrow(posdf))
emotion = rep(c("anxiety", "anger", "sadness", "positive"), times = c(nrow(anxdf),nrow(angerdf),nrow(saddf),nrow(posdf))) 
countrydf = cbind(rbind(anxdf, angerdf, saddf, posdf), emotion, country)
alldf = rbind(alldf, countrydf)
```

<!-- For tweets that contain emotional terms, the plot only shows terms not included among the most frequent in total tweet volume. -->

```{r, plots baseline spain,  fig.width=12, fig.height=9, cache=T, fig.cap ="**Figure S2.5. Most frequent terms in tweets per category during the baseline in Spain.** Percentage of tweets with the term."}
overlap = TRUE
ymax=10.5
ptot <- ggplot(totalbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="gray") +
  geom_text(aes(label = term, y=freq+0.2*nchar(as.character(term))/2), vjust = 0.5) +  ylim(0,ymax)+ylab("% total tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank()) 

panx <- ggplot(anxbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="orange") +
  geom_text(aes(label = term, y=freq+0.2*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% anxiety tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

psad <- ggplot(sadbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="blue") +
  geom_text(aes(label = term, y=freq+0.15*nchar(as.character(term))), vjust = 0.5)+ ylim(0,ymax)+ ylab("% sadness tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

pang <- ggplot(angerbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="red") +
  geom_text(aes(label = term, y=freq+0.2*nchar(as.character(term))/2), vjust = 0.5)+ylim(0,ymax)+ ylab("% anger tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ppos <- ggplot(posbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="cyan") +
  geom_text(aes(label = term, y=freq+0.2*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% positive tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ptot | panx | psad | pang |  ppos
```

\newpage

<!-- For tweets that contain emotional terms, the plot only shows terms not included among the most frequent in total tweet volume. -->

```{r, plots outbreak spain,  fig.width=12, fig.height=9, cache=T, fig.cap ="**Figure S2.6. Most frequent terms in tweets per category during the 5 weeks after the outbreak in Spain.** Percentage of tweets with the term."}
overlap = TRUE
ymax = 10
ptot <- ggplot(totaldf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="gray") +
  geom_text(aes(label = term, y=freq+0.2*nchar(as.character(term))/2), vjust = 0.5) +  ylim(0,ymax)+ylab("% total tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank()) 

panx <- ggplot(anxdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="orange") +
  geom_text(aes(label = term, y=freq+0.1*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% anxiety tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

psad <- ggplot(saddf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="blue") +
  geom_text(aes(label = term, y=freq+0.1*nchar(as.character(term))), vjust = 0.5)+ ylim(0,ymax)+ ylab("% sadness tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

pang <- ggplot(angerdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="red") +
  geom_text(aes(label = term, y=freq+0.1*nchar(as.character(term))/2), vjust = 0.5)+ylim(0,ymax)+ ylab("% anger tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ppos <- ggplot(posdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="cyan") +
  geom_text(aes(label = term, y=freq+0.1*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% positive tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())
ptot | panx | psad | pang | ppos
```

<!-- ### Difference: Volume of tweets with the term during baseline vs. 5 outbreak weeks -->

<!-- What increases, what decreases after the outbreak? Only words that are among the 60 most frequent during baseline are shown. New terms are shown in the second row. -->

<!-- ```{r, plots difference spain,  fig.width=12, fig.height=7} -->
<!-- ymax = 8; ymin = -2 -->

<!-- ptot <- ggplot(filter(totaldiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="grey") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.2*nchar(as.character(term))/2),vjust = 0.5)+ ylim(ymin,ymax)+ ylab("total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ ggtitle("Outbreak vs. baseline") -->

<!-- panx <- ggplot(filter(anxdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="orange") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.2*nchar(as.character(term))/2),vjust = 0.5)+ ylim(ymin,ymax)+ ylab("anxiety tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank()) -->
<!-- psad <- ggplot(filter(saddiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="blue") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.2*nchar(as.character(term))), vjust = 0.5)+  ylim(ymin,ymax)+ ylab("sadness vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- pang <- ggplot(filter(angerdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="red") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.2*nchar(as.character(term))/2), vjust = 0.5)+ ylim(ymin,ymax)+ ylab("anger vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- ppos <- ggplot(filter(posdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="cyan") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.2*nchar(as.character(term))/2), vjust = 0.5)+  ylim(ymin,ymax)+ ylab("positive vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- ptot | panx | psad | pang |ppos -->
<!-- ``` -->


<!-- ```{r  fig.width=12, fig.height=5} -->
<!-- #plots with new terms -->
<!-- ymax = 6 -->
<!-- newtot <- ggplot(filter(totaldiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="grey")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+ ylab("anxiety tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("New terms during outbreak") -->
<!-- newanx <- ggplot(filter(anxdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="orange")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.03*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+ ylab("anxiety tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank()) -->
<!-- newsad <- ggplot(filter(saddiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="blue")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.01*nchar(as.character(term))/2),vjust = 0.5)+  ylim(0,ymax)+ ylab("sadness tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newang <- ggplot(filter(angerdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="red")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.01*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+  ylab("anger tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newpos <- ggplot(filter(posdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="cyan")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.01*nchar(as.character(term))/2),vjust = 0.5)+  ylim(0,ymax)+ ylab("positive tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newtot | newanx | newsad | newang |  newpos -->
<!-- ``` -->


```{r, spanish liwc terms, cache = TRUE}
#adapt wordlist dfs for a table
angb = make_df_for_tab(angerbasedf, "baseline")
anxb = make_df_for_tab(anxbasedf, "baseline")
sadb = make_df_for_tab(sadbasedf, "baseline")
posb = make_df_for_tab(posbasedf, "baseline")
ang = make_df_for_tab(angerdf, "outbreak")
anx = make_df_for_tab(anxdf, "outbreak")
sad = make_df_for_tab(saddf, "outbreak")
pos = make_df_for_tab(posdf, "outbreak")

#get and preprocess LIWC dictionaries
#directory with the dictionary files
dirdictionaries <- '~/Documents/dictionaries/'
#load dictionaries and delete the *, because it means zero or more matches of the previous letter. $ at end of each liwc term without * searches only for words exactly ending like the liwc term 
dictang <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'spanish2007_anger.txt'))$V1, "$"), "\\*\\$", "")
dictanx <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'spanish2007_anx.txt'))$V1, "$"), "\\*\\$", "")
dictsad <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'spanish2007_sad.txt'))$V1, "$"), "\\*\\$", "")
dictpos <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'spanish2007_posemo.txt'))$V1, "$"), "\\*\\$", "")

# ## add "[\\s-]|^" at the start of each word to also include matches that include "-" or a white space (e.g. self-isolating, tut mir leid), and so that words have to start with the search term otherwise (^) (so "ok" does not match with "book")
dictang = paste0("[\\s-]", dictang, "|^", dictang)
dictanx = paste0("[\\s-]", dictanx, "|^", dictanx)
dictsad = paste0("[\\s-]", dictsad, "|^", dictsad)
dictpos = paste0("[\\s-]", dictpos, "|^", dictpos)

#add escape character to smiley brackets and colon
dictpos <- str_replace_all(dictpos, "\\(", "\\\\\\(")
dictpos <- str_replace_all(dictpos, "\\)", "\\\\\\)")
dictpos <- str_replace_all(dictpos, "\\:", "\\\\\\:")

# filter out LIWC terms
angbliwc =  angb %>% 
  filter(str_detect(Term, paste(dictang, collapse = "|")))
anxbliwc =  anxb %>% 
  filter(str_detect(Term, paste(dictanx, collapse = "|")))
sadbliwc =  sadb %>% 
  filter(str_detect(Term, paste(dictsad, collapse = "|")))
posbliwc =  posb %>% 
  filter(str_detect(Term, paste(dictpos, collapse = "|")))
angliwc =  ang %>% 
  filter(str_detect(Term, paste(dictang, collapse = "|")))
anxliwc =  anx %>% 
  filter(str_detect(Term, paste(dictanx, collapse = "|")))
sadliwc =  sad %>% 
  filter(str_detect(Term, paste(dictsad, collapse = "|")))
posliwc =  pos %>% 
  filter(str_detect(Term, paste(dictpos, collapse = "|")))

liwc = rbind(angbliwc, anxbliwc, sadbliwc, posbliwc, angliwc, anxliwc, sadliwc, posliwc) %>% 
  mutate(Emotion = recode(Emotion, pos = "positive", anx = "anxiety", sad = "sadness", ang = "anger")) %>% 
  relocate(c("Period", "Emotion"), .before="Term")

liwc_esp = liwc
```

**Table S2.3. Frequency of LIWC terms among the 60 most frequent terms in Spain per time period and emotion.**

```{r, out.width=0.5}
knitr::kable(liwc,
          digits = 2,
          format="html",
           table.attr = "style='width:50%;'") %>% 
  kable_classic(full_width=T, position = "left",  font_size = 13) 
```

```{r}
mins =round((c(min(anxb$Percent), min(anx$Percent), min(angb$Percent), min(ang$Percent), min(sadb$Percent), min(sad$Percent), min(posb$Percent), min(pos$Percent))),2)
lowest_percent = paste(mins[1], "and", mins[2], "for anxiety,", mins[3], "and", mins[4], "for anger,",mins[5], "and", mins[6], "for sadness, and",mins[7], "and", mins[8], "for positive emotions")
```

**Observations**

* In Spanish, the word "shame" (verguenza) seems to be an important term in the anxiety dictionary, both during baseline and post-covid outbreak.
* The word government was used more often after the outbreak, in particular in anxiety and anger tweets.
* The words house and state was generally used more often after the outbreak.

*LIWC words*: 

__No LIWC term was present in more than `r round(max(liwc$Percent), digits = 2)`% of tweets expressing a particular emotion. Percent of tweets containing all LIWC terms not listed in Table S2.3. during baseline and outbreak, respectively, were below `r lowest_percent` (percentages for words at rank 60 per dictionary and time period).__

* Anxiety: 
    - baseline: falta (missing), estrés (stress), depresión, shame (verguenza)
    - outbreak: falta, estrés, miedo (fear), panico, shame (verguenza)
* Sadness: daño (hurt), extraño/extraño (I miss, he/she/it misses), inutil (useless)
* Anger: ridículo (only baseline), coño (idiot), puta, mierda, mal, proteccion (defense)
* Positive: 
  - Baseline: smiley, espero (hope), facil (easy), super
  - Outbreak ánimo (encourage, encouragement or mood)

*Topics*: 

* Baseline: Politics (partido, PSOE, gobierno)
* Outbreak: Gobierno ranks very highly (especially in anxiety and anger tweets), crisis, cuarentena, casa (house) and gente/personas (people/persons) increased, queda (stay) 
    - Anxiety: medidas (measures), cuarentena, información, estado (de alarma; emergency state) increased particularly, presión (pressure), sanidad/salud (health)
    - Sad: cuarentena, salir (go out), casos (cases), pandemia, espero, less partido
    - Anger: medidas (measures),  more protección, less partido (political party), salir (go out)
    - Positive: mequedoencasa, medidas, familia, equipo, animo (encouragement), abrazo

\newpage

### Italian: Italy

```{r, prep data Italy, cache=T}
overlap = TRUE
country <- "italy"
totalbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_total_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

angerbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anger_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq))

sadbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_sad_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>%filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq))

anxbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anxiety_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>%filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

posbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_positive_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

# 5 week after outbreak
totaldf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_total_20200222-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  # exchange emojji picture for description
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# difference to baseline
totaldiffdf = totalbasedf %>%  select(-Type) %>%
  right_join(totaldf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))

angerdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anger_20200222-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
    mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# add columns for terms that exist in total and anger (36) and calculate difference
# angerdiffdf = angerdf %>%  select(-Type) %>%
#   inner_join(angerdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
angerdiffdf = angerbasedf %>%  select(-Type) %>%
  right_join(angerdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
#keep only terms that are not in totaldf
if(overlap==FALSE){angerdf = angerdf %>%
  filter(!term %in% totaldf$term)}

saddf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_sad_20200222-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# saddiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(saddf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
saddiffdf = sadbasedf %>%  select(-Type) %>%
  right_join(saddf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){saddf = saddf %>%
  filter(!term %in% totaldf$term)}

anxdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anxiety_20200222-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
   mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# anxdiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(anxdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
anxdiffdf = anxbasedf %>%  select(-Type) %>%
  right_join(anxdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){anxdf = anxdf %>%
  filter(!term %in% totaldf$term)}

posdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_positive_20200222-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
   mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# posdiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(posdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
posdiffdf = posbasedf %>%  select(-Type) %>%
  right_join(posdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){posdf = posdf %>%
  filter(!term %in% totaldf$term)}

### how many terms occur in more than 1% of the tweets (during outbreak)? 
country = rep(country, nrow(anxdf)+nrow(angerdf)+nrow(saddf)+nrow(posdf))
emotion = rep(c("anxiety", "anger", "sadness", "positive"), times = c(nrow(anxdf),nrow(angerdf),nrow(saddf),nrow(posdf))) 
countrydf = cbind(rbind(anxdf, angerdf, saddf, posdf), emotion, country)
alldf = rbind(alldf, countrydf)
```



<!-- For tweets that contain emotional terms, the plot only shows terms not included among the most frequent in total tweet volume. -->

```{r, plots baseline Italy,  fig.width=12, fig.height=9, cache=T, fig.cap ="**Figure S2.7. Most frequent terms in tweets per category during the baseline in Italy.** Percentage of tweets with the term."}
overlap = TRUE
ymax=3.5
ptot <- ggplot(totalbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="gray") +
  geom_text(aes(label = term, y=freq+0.02*nchar(as.character(term))/2), vjust = 0.5) +  ylim(0,ymax)+ylab("% total tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank()) 

panx <- ggplot(anxbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="orange") +
  geom_text(aes(label = term, y=freq+0.02*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% anxiety tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

psad <- ggplot(sadbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="blue") +
  geom_text(aes(label = term, y=freq+0.015*nchar(as.character(term))), vjust = 0.5)+ ylim(0,ymax)+ ylab("% sadness tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

pang <- ggplot(angerbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="red") +
  geom_text(aes(label = term, y=freq+0.02*nchar(as.character(term))/2), vjust = 0.5)+ylim(0,ymax)+ ylab("% anger tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ppos <- ggplot(posbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="cyan") +
  geom_text(aes(label = term, y=freq+0.02*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% positive tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ptot | panx | psad | pang |  ppos
```


\newpage



<!-- For tweets that contain emotional terms, the plot only shows terms not included among the most frequent in total tweet volume. -->

```{r, plots outbreak Italy,  fig.width=12, fig.height=9, cache = T, fig.cap ="**Figure S2.8. Most frequent terms in tweets per category during the 5 weeks after the outbreak in Italy.** Percentage of tweets with the term."}
overlap = TRUE
ymax = 15
ptot <- ggplot(totaldf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="gray") +
  geom_text(aes(label = term, y=freq+0.3*nchar(as.character(term))/2), vjust = 0.5) +  ylim(0,ymax)+ylab("% total tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank()) 

panx <- ggplot(anxdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="orange") +
  geom_text(aes(label = term, y=freq+0.3*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% anxiety tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

psad <- ggplot(saddf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="blue") +
  geom_text(aes(label = term, y=freq+0.2*nchar(as.character(term))), vjust = 0.5)+ ylim(0,ymax)+ ylab("% sadness tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

pang <- ggplot(angerdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="red") +
  geom_text(aes(label = term, y=freq+0.3*nchar(as.character(term))/2), vjust = 0.5)+ylim(0,ymax)+ ylab("% anger tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ppos <- ggplot(posdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="cyan") +
  geom_text(aes(label = term, y=freq+0.2*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% positive tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())
ptot | panx | psad | pang | ppos
```

<!-- ### Difference: Volume of tweets with the term during baseline vs. 5 outbreak weeks -->

<!-- What increases, what decreases after the outbreak? Only words that are among the 60 most frequent during baseline are shown. New terms are shown in the second row. -->

<!-- ```{r, plots difference Italy,  fig.width=12, fig.height=7} -->
<!-- ymax = 3; ymin = -0.7 -->

<!-- ptot <- ggplot(filter(totaldiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="grey") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.1*nchar(as.character(term))/2),vjust = 0.5)+ ylim(ymin,ymax)+ ylab("total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ ggtitle("Outbreak vs. baseline") -->

<!-- panx <- ggplot(filter(anxdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="orange") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.1*nchar(as.character(term))/2),vjust = 0.5)+ ylim(ymin,ymax)+ ylab("anxiety tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank()) -->
<!-- psad <- ggplot(filter(saddiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="blue") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.05*nchar(as.character(term))), vjust = 0.5)+  ylim(ymin,ymax)+ ylab("sadness vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- pang <- ggplot(filter(angerdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="red") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.1*nchar(as.character(term))/2), vjust = 0.5)+ ylim(ymin,ymax)+ ylab("anger vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- ppos <- ggplot(filter(posdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="cyan") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.1*nchar(as.character(term))/2), vjust = 0.5)+  ylim(ymin,ymax)+ ylab("positive vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- ptot | panx | psad | pang |ppos -->
<!-- ``` -->


<!-- ```{r  fig.width=12, fig.height=5} -->
<!-- #plots with new terms -->
<!-- ymax = 14.5 -->
<!-- newtot <- ggplot(filter(totaldiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="grey")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+ ylab("anxiety tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("New terms during outbreak") -->
<!-- newanx <- ggplot(filter(anxdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="orange")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+ ylab("anxiety tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank()) -->
<!-- newsad <- ggplot(filter(saddiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="blue")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+  ylim(0,ymax)+ ylab("sadness tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newang <- ggplot(filter(angerdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="red")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+  ylab("anger tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newpos <- ggplot(filter(posdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="cyan")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+  ylim(0,ymax)+ ylab("positive tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newtot | newanx | newsad | newang |  newpos -->
<!-- ``` -->

```{r, italian liwc terms, cache = TRUE}
#adapt wordlist dfs for a table
angb = make_df_for_tab(angerbasedf, "baseline")
anxb = make_df_for_tab(anxbasedf, "baseline")
sadb = make_df_for_tab(sadbasedf, "baseline")
posb = make_df_for_tab(posbasedf, "baseline")
ang = make_df_for_tab(angerdf, "outbreak")
anx = make_df_for_tab(anxdf, "outbreak")
sad = make_df_for_tab(saddf, "outbreak")
pos = make_df_for_tab(posdf, "outbreak")

#get and preprocess LIWC dictionaries
#directory with the dictionary files
dirdictionaries <- '~/Documents/dictionaries/'

#load dictionaries and delete the *, because it means zero or more matches of the previous letter. $ at end of each liwc term without * searches only for words exactly ending like the liwc term 
dictang <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'italian2007_anger.txt'))$V1, "$"), "\\*\\$", "")
dictanx <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'italian2007_anx.txt'))$V1, "$"), "\\*\\$", "")
dictsad <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'italian2007_sad.txt'))$V1, "$"), "\\*\\$", "")
dictpos <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'italian2007_posemo.txt'))$V1, "$"), "\\*\\$", "")

# ## add "[\\s-]|^" at the start of each word to also include matches that include "-" or a white space (e.g. self-isolating, tut mir leid), and so that words have to start with the search term otherwise (^) (so "ok" does not match with "book")
dictang = paste0("[\\s-]", dictang, "|^", dictang)
dictanx = paste0("[\\s-]", dictanx, "|^", dictanx)
dictsad = paste0("[\\s-]", dictsad, "|^", dictsad)
dictpos = paste0("[\\s-]", dictpos, "|^", dictpos)

# filter out LIWC terms
angbliwc =  angb %>% 
  filter(str_detect(Term, paste(dictang, collapse = "|")))
anxbliwc =  anxb %>% 
  filter(str_detect(Term, paste(dictanx, collapse = "|")))
sadbliwc =  sadb %>% 
  filter(str_detect(Term, paste(dictsad, collapse = "|")))
posbliwc =  posb %>% 
  filter(str_detect(Term, paste(dictpos, collapse = "|")))
angliwc =  ang %>% 
  filter(str_detect(Term, paste(dictang, collapse = "|")))
anxliwc =  anx %>% 
  filter(str_detect(Term, paste(dictanx, collapse = "|")))
sadliwc =  sad %>% 
  filter(str_detect(Term, paste(dictsad, collapse = "|")))
posliwc =  pos %>% 
  filter(str_detect(Term, paste(dictpos, collapse = "|")))

liwc = rbind(angbliwc, anxbliwc, sadbliwc, posbliwc, angliwc, anxliwc, sadliwc, posliwc) %>% 
  mutate(Emotion = recode(Emotion, pos = "positive", anx = "anxiety", sad = "sadness", ang = "anger")) %>% 
  relocate(c("Period", "Emotion"), .before="Term")

liwc_ita = liwc

#this is emty - no LIWC words among the most frequent. Report least frequent words per dictionary instead, all LIWC terms less than that.  
```

```{r}
mins =round((c(min(anxb$Percent), min(anx$Percent), min(angb$Percent), min(ang$Percent), min(sadb$Percent), min(sad$Percent), min(posb$Percent), min(pos$Percent))),2)
lowest_percent = paste(mins[1], "and", mins[2], "for anxiety,", mins[3], "and", mins[4], "for anger,",mins[5], "and", mins[6], "for sadness, and",mins[7], "and", mins[8], "for positive emotions")
```

**Observations**

*LIWC words*: 

__The 60 most frequent terms do not contain any of the LIWC terms, so all LIWC terms during baseline and outbreak, respectively, were below `r lowest_percent` (percentages for words at rank 60 per dictionary and time period).__

*Topics*: 

* Baseline: Politics - (Matteo) Salvini, Roma, PD (Partito Democratico)
* Outbreak:  Different forms of "to be"  (stanno, sto, storia), gente (people) in all emotional tweets, quarantena
    - Anxiety: contagion, stay (stare), emergency (emergenza), casi (cases), salute (health)
    - Sadness: Lombardia, quarantine, cazzo (fuck), stare (stay), casi (cases), uscire (going out)
    - Anger: cazzo (fuck)
    - Positive: stare (stay), abbraccio (hug), spero (hope)
* The name of Giuseppe Conte was more frequent in negative emotional tweets during the outbreak.
* The word stay (stare/stiamo) was more frequent in anxiety, sad and positive tweets during the outbreak.



\newpage

### Dutch: Netherlands

```{r, prep data netherlands, cache=T}
overlap = TRUE
country <- "netherlands"
totalbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_total_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

angerbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anger_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq))

sadbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_sad_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>%filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq))

anxbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anxiety_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>%filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

posbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_positive_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

# 5 week after outbreak
totaldf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_total_20200304-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  # exchange emojji picture for description
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# difference to baseline
totaldiffdf = totalbasedf %>%  select(-Type) %>%
  right_join(totaldf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))

angerdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anger_20200304-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
    mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# add columns for terms that exist in total and anger (36) and calculate difference
# angerdiffdf = angerdf %>%  select(-Type) %>%
#   inner_join(angerdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
angerdiffdf = angerbasedf %>%  select(-Type) %>%
  right_join(angerdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
#keep only terms that are not in totaldf
if(overlap==FALSE){angerdf = angerdf %>%
  filter(!term %in% totaldf$term)}

saddf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_sad_20200304-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# saddiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(saddf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
saddiffdf = sadbasedf %>%  select(-Type) %>%
  right_join(saddf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){saddf = saddf %>%
  filter(!term %in% totaldf$term)}

anxdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anxiety_20200304-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
   mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# anxdiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(anxdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
anxdiffdf = anxbasedf %>%  select(-Type) %>%
  right_join(anxdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){anxdf = anxdf %>%
  filter(!term %in% totaldf$term)}

posdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_positive_20200304-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
   mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# posdiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(posdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
posdiffdf = posbasedf %>%  select(-Type) %>%
  right_join(posdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){posdf = posdf %>%
  filter(!term %in% totaldf$term)}

### how many terms occur in more than 1% of the tweets (during outbreak)?
country = rep(country, nrow(anxdf)+nrow(angerdf)+nrow(saddf)+nrow(posdf)) 
emotion = rep(c("anxiety", "anger", "sadness", "positive"), times = c(nrow(anxdf),nrow(angerdf),nrow(saddf),nrow(posdf)))
countrydf = cbind(rbind(anxdf, angerdf, saddf, posdf), emotion, country)
alldf = rbind(alldf, countrydf)
```



<!-- For tweets that contain emotional terms, the plot only shows terms not included among the most frequent in total tweet volume. -->

```{r, plots baseline netherlands,  fig.width=12, fig.height=9, cache = T, fig.cap ="**Figure S2.9. Most frequent terms in tweets per category during the baseline in the Netherlands.** Percentage of tweets with the term."}
overlap = TRUE
ymax=8
ptot <- ggplot(totalbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="gray") +
  geom_text(aes(label = term, y=freq+0.02*nchar(as.character(term))/2), vjust = 0.5) +  ylim(0,ymax)+ylab("% total tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank()) 

panx <- ggplot(anxbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="orange") +
  geom_text(aes(label = term, y=freq+0.01*nchar(as.character(term))/2-0.25), vjust = 0.5)+ ylim(0,ymax)+ ylab("% anxiety tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

psad <- ggplot(sadbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="blue") +
  geom_text(aes(label = term, y=freq+0.015*nchar(as.character(term))), vjust = 0.5)+ ylim(0,ymax)+ ylab("% sadness tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

pang <- ggplot(angerbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="red") +
  geom_text(aes(label = term, y=freq+0.02*nchar(as.character(term))/2), vjust = 0.5)+ylim(0,ymax)+ ylab("% anger tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ppos <- ggplot(posbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="cyan") +
  geom_text(aes(label = term, y=freq+0.02*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% positive tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ptot | panx | psad | pang |  ppos
```

\newpage



<!-- For tweets that contain emotional terms, the plot only shows terms not included among the most frequent in total tweet volume. -->

```{r, plots outbreak netherlands,  fig.width=12, fig.height=9, cache=T, fig.cap ="**Figure S2.10. Most frequent terms in tweets per category during the 5 weeks after the outbreak in the Netherlands.** Percentage of tweets with the term."}
overlap = TRUE
ymax = 12
ptot <- ggplot(totaldf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="gray") +
  geom_text(aes(label = term, y=freq+0.2*nchar(as.character(term))/2), vjust = 0.5) +  ylim(0,ymax)+ylab("% total tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank()) 

panx <- ggplot(anxdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="orange") +
  geom_text(aes(label = term, y=freq+0.2*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% anxiety tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

psad <- ggplot(saddf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="blue") +
  geom_text(aes(label = term, y=freq+0.08*nchar(as.character(term))), vjust = 0.5)+ ylim(0,ymax)+ ylab("% sadness tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

pang <- ggplot(angerdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="red") +
  geom_text(aes(label = term, y=freq+0.15*nchar(as.character(term))/2), vjust = 0.5)+ylim(0,ymax)+ ylab("% anger tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ppos <- ggplot(posdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="cyan") +
  geom_text(aes(label = term, y=freq+0.15*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% positive tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())
ptot | panx | psad | pang | ppos
```

<!-- ### Difference: Volume of tweets with the term during baseline vs. 5 outbreak weeks -->

<!-- What increases, what decreases after the outbreak? Only words that are among the 60 most frequent during baseline are shown. New terms are shown in the second row. -->

<!-- ```{r, plots difference netherlands,  fig.width=12, fig.height=7} -->
<!-- ymax = 5; ymin = -2 -->

<!-- ptot <- ggplot(filter(totaldiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="grey") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.1*nchar(as.character(term))/2),vjust = 0.5)+ ylim(ymin,ymax)+ ylab("total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ ggtitle("Outbreak vs. baseline") -->

<!-- panx <- ggplot(filter(anxdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="orange") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.1*nchar(as.character(term))/2),vjust = 0.5)+ ylim(ymin,ymax)+ ylab("anxiety tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank()) -->
<!-- psad <- ggplot(filter(saddiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="blue") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.05*nchar(as.character(term))), vjust = 0.5)+  ylim(ymin,ymax)+ ylab("sadness vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- pang <- ggplot(filter(angerdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="red") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.1*nchar(as.character(term))/2), vjust = 0.5)+ ylim(ymin,ymax)+ ylab("anger vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- ppos <- ggplot(filter(posdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="cyan") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.1*nchar(as.character(term))/2), vjust = 0.5)+  ylim(ymin,ymax)+ ylab("positive vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- ptot | panx | psad | pang |ppos -->
<!-- ``` -->


<!-- ```{r  fig.width=12, fig.height=5} -->
<!-- #plots with new terms -->
<!-- ymax = 14.5 -->
<!-- newtot <- ggplot(filter(totaldiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="grey")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+ ylab("anxiety tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("New terms during outbreak") -->
<!-- newanx <- ggplot(filter(anxdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="orange")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+ ylab("anxiety tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank()) -->
<!-- newsad <- ggplot(filter(saddiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="blue")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+  ylim(0,ymax)+ ylab("sadness tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newang <- ggplot(filter(angerdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="red")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+  ylab("anger tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newpos <- ggplot(filter(posdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="cyan")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+  ylim(0,ymax)+ ylab("positive tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newtot | newanx | newsad | newang |  newpos -->
<!-- ``` -->


```{r, dutch liwc terms, cache = TRUE}
#adapt wordlist dfs for a table
angb = make_df_for_tab(angerbasedf, "baseline")
anxb = make_df_for_tab(anxbasedf, "baseline")
sadb = make_df_for_tab(sadbasedf, "baseline")
posb = make_df_for_tab(posbasedf, "baseline")
ang = make_df_for_tab(angerdf, "outbreak")
anx = make_df_for_tab(anxdf, "outbreak")
sad = make_df_for_tab(saddf, "outbreak")
pos = make_df_for_tab(posdf, "outbreak")

#get and preprocess LIWC dictionaries
#directory with the dictionary files
dirdictionaries <- '~/Documents/dictionaries/'
#load dictionaries and delete the *, because it means zero or more matches of the previous letter. $ at end of each liwc term without * searches only for words exactly ending like the liwc term 
dictang <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'dutch2007_anger.txt'))$V1, "$"), "\\*\\$", "")
dictanx <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'dutch2007_anx.txt'))$V1, "$"), "\\*\\$", "")
dictsad <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'dutch2007_sad.txt'))$V1, "$"), "\\*\\$", "")
dictpos <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'dutch2007_posemo.txt'))$V1, "$"), "\\*\\$", "")

# ## add "[\\s-]|^" at the start of each word to also include matches that include "-" or a white space (e.g. self-isolating, tut mir leid), and so that words have to start with the search term otherwise (^) (so "ok" does not match with "book")
dictang = paste0("[\\s-]", dictang, "|^", dictang)
dictanx = paste0("[\\s-]", dictanx, "|^", dictanx)
dictsad = paste0("[\\s-]", dictsad, "|^", dictsad)
dictpos = paste0("[\\s-]", dictpos, "|^", dictpos)

# filter out LIWC terms
angbliwc =  angb %>% 
  filter(str_detect(Term, paste(dictang, collapse = "|")))
anxbliwc =  anxb %>% 
  filter(str_detect(Term, paste(dictanx, collapse = "|")))
sadbliwc =  sadb %>% 
  filter(str_detect(Term, paste(dictsad, collapse = "|")))
posbliwc =  posb %>% 
  filter(str_detect(Term, paste(dictpos, collapse = "|")))
angliwc =  ang %>% 
  filter(str_detect(Term, paste(dictang, collapse = "|")))
anxliwc =  anx %>% 
  filter(str_detect(Term, paste(dictanx, collapse = "|")))
sadliwc =  sad %>% 
  filter(str_detect(Term, paste(dictsad, collapse = "|")))
posliwc =  pos %>% 
  filter(str_detect(Term, paste(dictpos, collapse = "|")))

liwc = rbind(angbliwc, anxbliwc, sadbliwc, posbliwc, angliwc, anxliwc, sadliwc, posliwc) %>% 
  mutate(Emotion = recode(Emotion, pos = "positive", anx = "anxiety", sad = "sadness", ang = "anger")) %>% 
  relocate(c("Period", "Emotion"), .before="Term")

liwc_ndl = liwc
```

    
```{r}
mins =round((c(min(anxb$Percent), min(anx$Percent), min(angb$Percent), min(ang$Percent), min(sadb$Percent), min(sad$Percent), min(posb$Percent), min(pos$Percent))),2)
lowest_percent = paste(mins[1], "and", mins[2], "for anxiety,", mins[3], "and", mins[4], "for anger,",mins[5], "and", mins[6], "for sadness, and",mins[7], "and", mins[8], "for positive emotions")
```


**Observations**

* In Dutch, the words "people", "all",  "tijd" (time) and "just" became more frequent during the outbreak
* The word "new year" became less frequent during the outbreak, simply because new years eve is not part of that time period, and Dutch Twitter apparently has a lot of new years wishes. The words new "nieuwe" and happy "gelukkig" are frequent as well. 

*LIWC words*: 

__Only one LIWC term  ("ongemak") was frequent enough (`r round(liwc$Percent, 2)`%) to be among the 60 most frequent terms. Percent of tweets containing all other LIWC terms during baseline and outbreak, respectively, were below `r lowest_percent` (percentages for words at rank 60 per dictionary and time period)__

__Words that might indicate a problem:__

* **"Ongemak" frequently occurs in the phrase "Please apologize for the inconvenience", which does not seem like an expression of anxiety. Yet, this term only occurs in `r round(liwc$Percent, 2)`% of all anxiety tweets, and can thus not have a strong influence on total anxiety scores.**

*Topics*: 

* Outbreak : coronavirus, thuis (at home), blijven (stay), nodig (necessary), elkaar (each other), sterkte (strength)
    - Anxiety: crisis, nodig (necessary), elkaar (each other)
    - Sad: Kinderen (kids)

    


\newpage


### French: France

```{r, prep data france, cache=T}
overlap = TRUE
country <- "france"
totalbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_total_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

angerbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anger_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq))

sadbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_sad_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>%filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq))

anxbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anxiety_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>%filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

posbasedf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_positive_2019.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term))%>%
  arrange(desc(freq))

# 5 week after outbreak
totaldf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_total_20200227-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  # exchange emojji picture for description
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# difference to baseline
totaldiffdf = totalbasedf %>%  select(-Type) %>%
  right_join(totaldf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))

angerdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anger_20200227-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
    mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# add columns for terms that exist in total and anger (36) and calculate difference
# angerdiffdf = angerdf %>%  select(-Type) %>%
#   inner_join(angerdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
angerdiffdf = angerbasedf %>%  select(-Type) %>%
  right_join(angerdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
#keep only terms that are not in totaldf
if(overlap==FALSE){angerdf = angerdf %>%
  filter(!term %in% totaldf$term)}

saddf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_sad_20200227-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
  mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# saddiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(saddf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
saddiffdf = sadbasedf %>%  select(-Type) %>%
  right_join(saddf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){saddf = saddf %>%
  filter(!term %in% totaldf$term)}

anxdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_anxiety_20200227-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
   mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# anxdiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(anxdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
anxdiffdf = anxbasedf %>%  select(-Type) %>%
  right_join(anxdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){anxdf = anxdf %>%
  filter(!term %in% totaldf$term)}

posdf <- read.csv(file=paste0("./data_wordcloudPeriods/wordcloud_",country,"_positive_20200227-5weeks.csv"), header=T, skip=6) %>%
  rename(term=as.character("Topic.Name"), freq = "Percentage.Volume" )%>% filter(term != "unrecognised_emoji") %>%
   mutate(term = makeemoji(term)) %>%
  arrange(desc(freq)) %>%
  mutate(term = tolower(term))
# posdiffdf = totaldf %>%  select(-Type) %>%
#   inner_join(posdf, by="term", suffix=c(".total", ".emo")) %>%
#   mutate(freq_diff = freq.emo-freq.total)
posdiffdf = posbasedf %>%  select(-Type) %>%
  right_join(posdf, by="term", suffix=c(".base", ".post")) %>%
  mutate(freq_diff = freq.post-freq.base, term = tolower(term))
if(overlap==FALSE){posdf = posdf %>%
  filter(!term %in% totaldf$term)}

### how many terms occur in more than 1% of the tweets (during outbreak)? 
country = rep(country, nrow(anxdf)+nrow(angerdf)+nrow(saddf)+nrow(posdf))
emotion = rep(c("anxiety", "anger", "sadness", "positive"), times = c(nrow(anxdf),nrow(angerdf),nrow(saddf),nrow(posdf))) 
countrydf = cbind(rbind(anxdf, angerdf, saddf, posdf), emotion, country)
alldf = rbind(alldf, countrydf)
```



<!-- For tweets that contain emotional terms, the plot only shows terms not included among the most frequent in total tweet volume. -->

```{r, plots baseline france,  fig.width=12, fig.height=9, cache=TRUE, fig.cap ="**Figure S2.11. Most frequent terms in tweets per category during the baseline in France.** Percentage of tweets with the term."}
overlap = TRUE
ymax=4
ptot <- ggplot(totalbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="gray") +
  geom_text(aes(label = term, y=freq+0.02*nchar(as.character(term))/2), vjust = 0.5) +  ylim(0,ymax)+ylab("% total tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank()) 

panx <- ggplot(anxbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="orange") +
  geom_text(aes(label = term, y=freq+0.01*nchar(as.character(term))/2-0.25), vjust = 0.5)+ ylim(0,ymax)+ ylab("% anxiety tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

psad <- ggplot(sadbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="blue") +
  geom_text(aes(label = term, y=freq+0.015*nchar(as.character(term))), vjust = 0.5)+ ylim(0,ymax)+ ylab("% sadness tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

pang <- ggplot(angerbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="red") +
  geom_text(aes(label = term, y=freq+0.02*nchar(as.character(term))/2), vjust = 0.5)+ylim(0,ymax)+ ylab("% anger tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ppos <- ggplot(posbasedf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="cyan") +
  geom_text(aes(label = term, y=freq+0.02*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% positive tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ptot | panx | psad | pang |  ppos
```

\newpage



<!-- For tweets that contain emotional terms, the plot only shows terms not included among the most frequent in total tweet volume. -->

```{r, plots outbreak france,  fig.width=12, fig.height=9, cache=TRUE, fig.cap ="**Figure S2.12. Most frequent terms in tweets per category during the 5 weeks after the outbreak in France.** Percentage of tweets with the term."}
overlap = TRUE
ymax = 5
ptot <- ggplot(totaldf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="gray") +
  geom_text(aes(label = term, y=freq+0.1*nchar(as.character(term))/2), vjust = 0.5) +  ylim(0,ymax)+ylab("% total tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank()) 

panx <- ggplot(anxdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="orange") +
  geom_text(aes(label = term, y=freq+0.1*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% anxiety tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

psad <- ggplot(saddf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="blue") +
  geom_text(aes(label = term, y=freq+0.05*nchar(as.character(term))), vjust = 0.5)+ ylim(0,ymax)+ ylab("% sadness tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

pang <- ggplot(angerdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="red") +
  geom_text(aes(label = term, y=freq+0.1*nchar(as.character(term))/2), vjust = 0.5)+ylim(0,ymax)+ ylab("% anger tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())

ppos <- ggplot(posdf, aes(x = reorder(term, freq), y = freq)) + geom_col(fill="cyan") +
  geom_text(aes(label = term, y=freq+0.08*nchar(as.character(term))/2), vjust = 0.5)+ ylim(0,ymax)+ ylab("% positive tweets with term")+
  coord_flip() + theme_bw() + theme(axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank())
ptot | panx | psad | pang | ppos
```

<!-- ### Difference: Volume of tweets with the term during baseline vs. 5 outbreak weeks -->

<!-- What increases, what decreases after the outbreak? Only words that are among the 60 most frequent during baseline are shown. New terms are shown in the second row. -->

<!-- ```{r, plots difference france,  fig.width=12, fig.height=7} -->
<!-- ymax = 2; ymin = -0.7 -->

<!-- ptot <- ggplot(filter(totaldiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="grey") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.1*nchar(as.character(term))/2),vjust = 0.5)+ ylim(ymin,ymax)+ ylab("total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ ggtitle("Outbreak vs. baseline") -->

<!-- panx <- ggplot(filter(anxdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="orange") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.1*nchar(as.character(term))/2),vjust = 0.5)+ ylim(ymin,ymax)+ ylab("anxiety tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank()) -->
<!-- psad <- ggplot(filter(saddiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="blue") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.05*nchar(as.character(term))), vjust = 0.5)+  ylim(ymin,ymax)+ ylab("sadness vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- pang <- ggplot(filter(angerdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="red") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.1*nchar(as.character(term))/2), vjust = 0.5)+ ylim(ymin,ymax)+ ylab("anger vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- ppos <- ggplot(filter(posdiffdf, !is.na(freq.base)), aes(x = reorder(term, freq_diff), y = freq_diff)) + geom_col(fill="cyan") + -->
<!--   geom_text(aes(label = term, y=freq_diff+0.1*nchar(as.character(term))/2), vjust = 0.5)+  ylim(ymin,ymax)+ ylab("positive vs. total tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- ptot | panx | psad | pang |ppos -->
<!-- ``` -->


<!-- ```{r  fig.width=12, fig.height=5} -->
<!-- #plots with new terms -->
<!-- ymax = 10 -->
<!-- newtot <- ggplot(filter(totaldiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="grey")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+ ylab("anxiety tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("New terms during outbreak") -->
<!-- newanx <- ggplot(filter(anxdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="orange")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+ ylab("anxiety tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank()) -->
<!-- newsad <- ggplot(filter(saddiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="blue")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+  ylim(0,ymax)+ ylab("sadness tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newang <- ggplot(filter(angerdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="red")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+ ylim(0,ymax)+  ylab("anger tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newpos <- ggplot(filter(posdiffdf, is.na(freq.base)), aes(x = reorder(term, freq.post), y=freq.post))+geom_col(fill="cyan")+ -->
<!--   geom_text(aes(label = term, y=freq.post+0.3*nchar(as.character(term))/2),vjust = 0.5)+  ylim(0,ymax)+ ylab("positive tweets")+ -->
<!--   coord_flip() + theme_bw() + theme(axis.title.y=element_blank(), -->
<!--                                     axis.text.y=element_blank(), -->
<!--                                     axis.ticks.y=element_blank())+ggtitle("") -->
<!-- newtot | newanx | newsad | newang |  newpos -->
<!-- ``` -->


```{r, French liwc terms, cache = TRUE}
#adapt wordlist dfs for a table
angb = make_df_for_tab(angerbasedf, "baseline")
anxb = make_df_for_tab(anxbasedf, "baseline")
sadb = make_df_for_tab(sadbasedf, "baseline")
posb = make_df_for_tab(posbasedf, "baseline")
ang = make_df_for_tab(angerdf, "outbreak")
anx = make_df_for_tab(anxdf, "outbreak")
sad = make_df_for_tab(saddf, "outbreak")
pos = make_df_for_tab(posdf, "outbreak")

#get and preprocess LIWC dictionaries
#directory with the dictionary files
dirdictionaries <- '~/Documents/dictionaries/'
#load dictionaries and delete the *, because it means zero or more matches of the previous letter. $ at end of each liwc term without * searches only for words exactly ending like the liwc term 
dictang <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'french2007_anger_no_noambiguous.txt'))$V1, "$"), "\\*\\$", "")
#no words excluded in french anger,  just stem* included for words for better overview rather than all declinations of words
dictanx <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'french2007_anx.txt'))$V1, "$"), "\\*\\$", "")
dictsad <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'french2007_sad.txt'))$V1, "$"), "\\*\\$", "")
dictpos <-  str_replace_all(paste0(read.table(paste0(dirdictionaries, 'french2007_posemo.txt'))$V1, "$"), "\\*\\$", "")

# ## add "[\\s-]|^" at the start of each word to also include matches that include "-" or a white space (e.g. self-isolating, tut mir leid), and so that words have to start with the search term otherwise (^) (so "ok" does not match with "book")
dictang = paste0("[\\s-]", dictang, "|^", dictang)
dictanx = paste0("[\\s-]", dictanx, "|^", dictanx)
dictsad = paste0("[\\s-]", dictsad, "|^", dictsad)
dictpos = paste0("[\\s-]", dictpos, "|^", dictpos)

# filter out LIWC terms
angbliwc =  angb %>% 
  filter(str_detect(Term, paste(dictang, collapse = "|")))
anxbliwc =  anxb %>% 
  filter(str_detect(Term, paste(dictanx, collapse = "|")))
sadbliwc =  sadb %>% 
  filter(str_detect(Term, paste(dictsad, collapse = "|")))
posbliwc =  posb %>% 
  filter(str_detect(Term, paste(dictpos, collapse = "|")))
angliwc =  ang %>% 
  filter(str_detect(Term, paste(dictang, collapse = "|")))
anxliwc =  anx %>% 
  filter(str_detect(Term, paste(dictanx, collapse = "|")))
sadliwc =  sad %>% 
  filter(str_detect(Term, paste(dictsad, collapse = "|")))
posliwc =  pos %>% 
  filter(str_detect(Term, paste(dictpos, collapse = "|")))

liwc = rbind(angbliwc, anxbliwc, sadbliwc, posbliwc, angliwc, anxliwc, sadliwc, posliwc) %>% 
  mutate(Emotion = recode(Emotion, pos = "positive", anx = "anxiety", sad = "sadness", ang = "anger")) %>% 
  relocate(c("Period", "Emotion"), .before="Term")
liwc_fr = liwc
```

**Table S2.4. Frequency of LIWC terms among the 60 most frequent terms in the France per time period and emotion.**

```{r, out.width=0.5}
knitr::kable(liwc,
          digits = 2,
          format="html",
           table.attr = "style='width:50%;'") %>% 
  kable_classic(full_width=T, position = "left", font_size = 13)
```

```{r}
mins =round((c(min(anxb$Percent), min(anx$Percent), min(angb$Percent), min(ang$Percent), min(sadb$Percent), min(sad$Percent), min(posb$Percent), min(pos$Percent))),2)
lowest_percent = paste(mins[1], "and", mins[2], "for anxiety,", mins[3], "and", mins[4], "for anger,",mins[5], "and", mins[6], "for sadness, and",mins[7], "and", mins[8], "for positive emotions")
```


**Observations**

*LIWC words*: 

__No LIWC term was present in more than `r round(max(liwc$Percent), digits = 2)`% of tweets expressing a particular emotion. Percent of tweets containing all LIWC terms not listed in Table S2.4. during baseline and outbreak, respectively, were below `r lowest_percent` (percentages for words at rank 60 per dictionary and time period).__

* Anxiety: In French, inquiète (worried), énerve (nervous/irritated), hésite (hesitate), lâche/lâcher (give me a break, or letting go?), dérange (disturb) are important words in the anxiety dictionary.
    - Outbreak: worried is more frequent
* Sadness: mal (wrong/bad) and deçu (disappointed) (in both time periods), echec (failure) and  défaite (defeat) (only during baseline) are important words in the sadness dictionary
* Anger: déteste (hate) is important in the angry dictionary and slighly less frequent during outbreak than during baseline. 
* Positive: no LIWC word among the most frequent

*Topics*: 
    
* Outbreak: cas (i.e. covid-19 cases), days, confinement, coronavirus/virus, pleading face emoji, never, life, laughing emoji were less frequent
    - Anxiety: health (santé), crisis, macron, the situation, the government, the cause of the situations
    - Sadness: pleading face emoji, mort (death), more crying emojis, crisis, situation
    - Anger: sortir (going outside)
    - Positive: pleading face emoji, less twinking emoji

Words that might indicate a problem: 

* McDo and Mcdonald in anxiety and positive tweets during the baseline, but without other terms that are part of French McDonald  ("C'est tout ce que j'aime", "Venez comme vous êtes"), and only in less than 1% of tweets. 

<!-- # Across all 6 countries -->

<!-- During the outbreak period:  -->

<!-- * Proportion of tweets that occur more often than in 3% of tweets: `r alldf %>% group_by(freq>3) %>%  summarise(pr = n()/nrow(alldf)) %>% slice(2) %>% select(pr)`. -->
<!-- * On average, words occur in `r median(alldf$freq)`% of tweets (median). -->
<!-- * The most frequent word is `r alldf[which(alldf$freq > 11.9), "term"]` and it occurs in `r max(alldf$freq)`% of tweets -->


\newpage

```{r, all liwc terms table}
#add all LIWC terms together, optionally show this as one table (uncomment the code below for this)
addcountry  = function(df, country){
  df = df %>% 
  mutate(Country = rep(country, nrow(df)))
  return(df)
  }
all_liwc= rbind(addcountry(liwc_uk, "UK"), addcountry(liwc_ger, "Germany"), addcountry(liwc_esp, "Spain"), addcountry(liwc_ita, "Italy"), addcountry(liwc_ndl, "Netherlands"), addcountry(liwc_fr, "France")) %>% 
  relocate(Country, .before = Period)
all_liwc$Percent= round(all_liwc$Percent, 2)

#the most common LIWC terms
all_liwc_ranked = all_liwc[order(all_liwc$Percent, decreasing=T),]

# write.csv(all_liwc, file='../output/Table_LIWCterms_60mostfrequent.csv', row.names = F)
# knitr::kable(all_liwc,
#           digits = 2,
#           format="html",
#            table.attr = "style='width:50%;'") %>%
#   kable_classic(full_width=T, position = "left",  font_size = 13)
```


## Summary and Discussion

Most words occur only in a small percentage of tweets. On average, terms occur in `r round(median(alldf$freq),2)`% (median) of tweets, with only  `r round(alldf %>% group_by(freq>3) %>%  summarise(pr = n()/nrow(alldf)) %>% slice(2) %>% select(pr),2)`% of terms occuring in more than 3% of tweets. The vast majority of these terms is not part of the LIWC dictionaries we used to assess emotion expression, but comprises generally frequent words like people, coronavirus, time, simply, normally, government etc. 

__The median frequency of LIWC terms is `r round(median(all_liwc$Percent), 2)`%, and the most frequent are `r all_liwc_ranked$Term[1]` with `r all_liwc_ranked$Percent[1]`% in tweets expressing anxiety, and `r all_liwc_ranked$Term[3]` with `r all_liwc_ranked$Percent[3]`% in tweets expressing positive emotions. Vergüenza, the Spanish word for shame, is often used in phrases like "what a shame" or "it's a shame that...". Given that this term is similarly often used during baseline and outbreak, it cannot explain the increases in anxiety we observe. Similarly, `r all_liwc_ranked$Term[3]` appears equally often during baseline and outbreak. All other terms occur in less than  `r all_liwc_ranked$Percent[5]`% of tweets expressing each particular emotion.__

Overall, these results suggests that no one word strongly misfired in response to a single event and thus had a strong effect on the change in each emotion measure. Instead, the count of emotional tweets per category reflects the sum of a broadly distributed higher occurence of emotion terms in general.

__Examples of other LIWC terms that made it into the top 60 in some of the countries are: (1) panic, fear, anxiety, pressure and stress in anxiety tweets; (2) isolating, disappointed, failure, tired, (I am) sorry, hurt, miss and useless in sadness tweets; (3) different swear words and mean, ridiculous, fight in anger tweets; and (4) smileys, happy, funny, heart, easy, beautiful, luck, and thank you in positive tweets.__ 

Regarding the topics of conversations in which emotional expressions occured, word frequencies reveal several trends that are visible across countries and emotions: First, the non-emotional words reveal a shift of topic discussed in emotional tweets from 2019 to the oubreak of COVID-19 from politics (typical terms: party, politician and country names, Brexit, government) to topics associated with the pandemic (coronavirus, virus, covid, measures, crisis etc) and the related everyday life changes. During the outbreak, all kinds of tweets more often mentioned people (and similar terms like children, family, staff, women), home, work, life, hope, and time (and similar terms like day, today, week). Second, words related to social distancing (quarantine, isolation, confinement, social, lockdown, stay at home, going out) often rank higher not only in sadness tweets, but also other emotional tweets, compared to tweets in general. That they also appear in positive terms, along with terms like home, books, team, family, hug etc. suggests that people noticed positive aspects of spending more time at home. 

Other topics that are more strongly linked to a particluar emotion are health, staying (at home), buying (of groceries), emergency, contagion and information for anxiety, as well as going out for sadness and anger, at home for sadness but also anxiety. The word (COVID-19) cases appears frequently in anxiety and sadness tweets in Spain, Italy, and France - which are all countries that the pandemic hit hard during the first outbreak period. Deaths are only mentioned in sadness tweets in France (otherwise also in French anger and Italian anxiety tweets). Altogether, topics in sadness tweets hint that changes to people's everyday life, rather than deaths, were a driving factor of increases in sadness terms. 





<!-- Case of Spain (for country comparisons): In the only country with a slight increase in anger, i.e. Spain, the terms government, state and measures rank high in anxiety and/or anger tweets, and anger tweets more often mention protection and going out. That these issues were often raised in a negative emotional context likely reflects that early pandemic-related decisions were made in a politically polarized context.  -->

```{r, results=FALSE}
# see the frequent terms
alldf %>% 
  #only terms that appear in more than 3 percent of tweets:
  filter(freq > 3) %>%
  group_by(term, emotion) %>%
  #average across countries
  summarise(min=min(freq),
            max=max(freq),
            mean=mean(freq)) %>%
  arrange(desc(mean))

```
```{r, results=FALSE}
#insert a string to check in which countries/emotions it occurs how often
alldf %>% 
  select(term, freq, emotion, country) %>%
  filter(str_detect(term, "ánimo")) #insert string here

#copy into deepl for each country to check for emotional words
alldf %>% filter(country=="spain") %>% select(term)
```